---
title: 'Assignment 1.1: Module 1 Exercises'
author: "Zachariah Freitas"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
defaultW <- getOption("warn") 
options(warn = -1) 
```

------------------------------------------------------------------------

## Chapter 1:

Approaching Forecasting (Page 23) #1-5

### Question 1.

Is the goal of this study descriptive or predictive?

#### Question CH1-Q1 Answer:

**Answer:** The goal of this study is to be descriptive.
To understand the impact that 9/11 had on long distance passenger travel behavior patterns.

### Question 2.

What is the forecast horizon to consider in this task?
Are next-month forecasts sufficient?

#### Question CH1-Q2 Answer:

```{r 911}
library(readxl)
library(zoo)
library(tidyverse)
library(fpp2) # Plot and Forecast Data
set.seed(506)


Sept11Travel <- read_excel("Sept11Travel.xls", 
    col_types = c("date", "numeric", "numeric", 
        "numeric"))


ggplot(Sept11Travel, aes(x = Month, y = `Air RPM (000s)`, group = 1)) +
  geom_line(size = 1, color = "black") +
  theme_classic()

# 
# ggplot(Sept11Travel, aes(x = Month, y = `Rail PM`, group = 1)) +
#   geom_line(size = 1, color = "green") +
#   theme_classic()
# 
# 
# ggplot(Sept11Travel, aes(x = Month, y = `VMT (billions)`, group = 1)) +
#   geom_line(size = 1, color = "black") +
#   theme_classic()

```

**Answer:** Impacts to 9/11 was felt for years after the attack.
You can visually see the impact in the chart above.
So looking at a minimum of several months to several years would be appropriate.

### Question 3.

What level of automation does this forecasting task require?
Consider the four questions related to automation.

#### Question CH1-Q3 Answer:

**Answer:** Given my answers below the level automation required for this task is relatively low.

1.  How many series need to be forecasted?\

    **Answer:** Only three, "Air RPM (000s)", "Rail PM", and "VMT (billions)".

2.  Is the forecasting an ongoing process or a one time event?\

    **Answer:** This forecast is a one time event.
    We hope.

3.  Which data and software will be available during the forecasting period?\

    **Answer:** A one time pull of Sept11Travel.xls data will be available and R will be used to analyze this data.
    With the following additional libraries, readxl, zoo, tidyverse, and fpp2.

4.  What forecasting expertise will be available **Answer:** Graduate level data scientist will be available for this task.

### Question 4.

What does the meaning of t = 1,2,3 in the Air series?
Which time period does t = 1 refer to?

#### Question CH1-Q4 Answer:

**Answer:** t = 1,2,3 is an index denoting the time period of interest, with this data t = monthly periods.
t = 1 is the fist period in the series, starting in January of 1990.

### Question 5.

What are the values for $y_1$, $y_2$, and $y_3$ in the Air series?

#### Question CH1-Q5 Answer:

**Answer:** $y_1$, $y_2$, and $y_3$ in the Air series, is a series of n values measured over n time periods, where $y_t$ denotes the value of the series at time period t.
For example, this data is a series of monthly Air RPM (000's), t =1,2,3,... denotes month 1, month 2, and month3; $y_1$, $y_2$, and $y_3$ denote the Air RPM (000's) on months 1, 2, and 3.

------------------------------------------------------------------------

## Chapter 2:

Time Series Data (Page 43) #3

### Question 3.

Shipments of Household Appliance: The file ApplianceShipments.xls contains the series of quarterly shipments (in millions of USD) of U.S. household appliances between 1985-1989.

a)  Create a well-formatted time plot of the data.

b)  Which of the four components (level, trend, seasonality, noise) seem to be present in this series?

#### Question CH2-Q3 Answer:

```{r appliance_shipments}
# Get data
ApplianceShipments <- read_excel("ApplianceShipments.xlsx")

# Convert Quarter character to date
ApplianceShipments$Quarter <-
  as.Date(
    as.yearqtr(
      ApplianceShipments$Quarter,
      format = "Q%q-%Y"
      ),
    # "frac=1" sets date to last day of quarter
    # frac = 1 
    )


# Create Time Series Object
MyTS <- ts(ApplianceShipments$Shipments, 
           start = c(1985, 1), 
           frequency = 4)


# Plot the Time Series
autoplot(MyTS) +
  labs(title = "Appliance Shipments Over Time",
       x = "Quarter",
       y = "Shipments") + 
  theme_minimal()


```

**Answer:** Looking at the chart above that we can see traces of all four components present in the series.
Some of the components are stronger than others, I would order them in terms of seasonality, trend as close contenders for strongest, followed by trend and then noise.

------------------------------------------------------------------------

## Chapter 3:

Performance Evaluation (Pages 67-68) #2 and 3

### Question 2.

Forecasting Shampoo Sales: The file ShampooSales.xls contains data on the monthly sales of a certain shampoo over a three-year period.
If the goal is forecasting sales in future months, which of the following steps should be taken?
(choose one or more below).

-   [x] Partition the data into training and validation periods.
-   [x] Examine time plots of the series and of model forecasts only for the training period.
-   [x] Look at MAPE and RMSE values for the training period.
-   [x] Look at MAPE and RMSE values for the validation period.
-   [x] Compute naïve forecasts. Page 1 of 2

#### Question CH3-Q2 Answer:
**Answer:** I added more models and naïve forecasts than asked for. The best model appears to be the Naïve: Drift Method followed by the Exponential and then the seasonal trend model. They beat all other naïve forecasts and models in RMSE and other metrics. See my results below.


```{r shampoo_model_comparison}
# Get Data
ShampooSales <- read_excel("ShampooSales.xlsx")

# Fix Date
ShampooSales$Month <- as.Date(ShampooSales$Month)

# Set Forcast periods
h=12

# Create Time Series Object
MyTS <- ts(ShampooSales$Shampoo_sales, 
           start = c(1995, 1), 
           frequency = 12)

# Split Data
train <- window(MyTS, 
                start = c(1995, 1), 
                end = c(1996, 12))


# Naïve Models
mean_fit <- meanf(train,h=h) # Naïve: Mean
naive_fit <- naive(train,h=h,level = 95) # Naïve: Last Value
drift_fit <- rwf(train,h=h,drift=TRUE) # Naïve: Drift Method
snaive_fit <- snaive(train,h=h) # Naïve: Seasonal Naïve

# Train Models
st_model <- tslm(train ~ trend + season)
st_fit <- forecast(st_model,h=h)
exp_model <- tslm(train ~ trend + season, lambda = 0)
exp_fit <- forecast(exp_model,h=h)
spline_model <- tslm(train ~ trend + I(trend^2) + I(trend^3) + season)
spine_fit <- forecast(spline_model,h=h)

# Plot Results
autoplot(window(MyTS, start=1995)) +
  autolayer(mean_fit, series="Naïve: Mean", PI=FALSE) +
  autolayer(drift_fit, series="Naïve: Drift Method", PI=FALSE) +
  autolayer(snaive_fit, series="Naïve: Seasonal Naïve", PI=FALSE) +
  autolayer(naive_fit, series="Naïve: Last Value", PI=FALSE) +
  autolayer(st_fit, series="Seasonal Trend", PI=FALSE) +
  autolayer(exp_fit, series="Exponential", PI=FALSE) +
  autolayer(spine_fit, series="Spline", PI=FALSE) +
  xlab("Month") + 
  ylab("Sales") +
  ggtitle("Forecasts for monthly shampoo sales") +
  theme_minimal() +
  guides(colour=guide_legend(title="Forecast"))
```

```{r train_vs_validation_results}
window_df <- window(MyTS, start=1997)
print(strrep("#", 80))
print("Naïve: Mean")
accuracy(mean_fit, window_df)

print(strrep("#", 80))
print("Naïve: Last Value")
accuracy(naive_fit, window_df)

print(strrep("#", 80))
print("Naïve: Drift Method")
accuracy(drift_fit, window_df)

print(strrep("#", 80))
print("Naïve: Seasonal Naïve")
accuracy(snaive_fit, window_df)

print(strrep("#", 80))
print("Seasonal Trend")
accuracy(st_fit, window_df)

print(strrep("#", 80))
print("Exponential")
accuracy(exp_fit, window_df)

print(strrep("#", 80))
print("Spline")
accuracy(spine_fit, window_df)

```

```{r rolling_cross_validation}
# This is here for my own edification.
e <- tsCV(train, rwf, drift=TRUE, h=1)
sqrt(mean(e^2, na.rm=TRUE))
sqrt(mean(residuals(rwf(train, drift=TRUE))^2, na.rm=TRUE))
```

### Question 3.

Performance on Training and Validation Data: Two different models were fit to the same time series.
The first 100 time periods were used for the training period and the last 12 periods were treated as a validation period.
Assume that both models make sense practically and fit the data reasonably well.
Below are the RMSE values for each of the models:

| Model   | Training Period | Validation Period |
|:--------|:---------------:|:-----------------:|
| Model A |       543       |        690        |
| Model B |       669       |        675        |

a)  Which model appears more useful for retrospectively describing the different components of this time series?
    Why?

b)  Which model appears to be more useful for forecasting purposes?
    Why?

#### Question CH3-Q3 Answer:

**Answer:**

a)  Model A appears more useful for retrospectively describing the different components of this time series because the root mean square error is smaller than that of model B suggesting a model fit is tighter and more reflective of what transpired during that time frame.

b)  Model B appears more useful for forecasting purposes because the root mean square error is smaller than that of model A.
    This suggests that model B generalizes better and is less over fit.
    

```{r undo_warnings, include=FALSE}
options(warn = defaultW)
```


