---
title: 'Assignment 1.1: Module 1 Exercises'
author: "Zachariah Freitas"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 1:

Approaching Forecasting (Page 23) #1-5 

### Question 1.
Is the goal of this study descriptive or predictive?

#### Question CH1-Q1 Answer:

### Question 2.

What is the forecast horizon to consider in this task?
Are next-month forecasts sufficient?

#### Question CH1-Q2 Answer:

### Question 3.

What level of automation does this forecasting task require? Consider the four questions related to automation.

#### Question CH1-Q3 Answer:

### Question 4.

What does the meaning of t = 1,2,3 in the Air series?
Which time period does t = 1 refer to?

#### Question CH1-Q4 Answer:

### Question 5.

What are the values for y1, y2, and y3 in the Air series?

#### Question CH1-Q5 Answer:

```{r cars}
summary(cars)
```

## Chapter 2:

Time Series Data (Page 43) #3

### Question 3.

Shipments of Household Appliance: The file ApplianceShipments.xls contains the series of quarterly shipments (in millions of USD) of U.S. household appliances between 1985-1989.

a)  Create a well-formatted time plot of the data.

b)  Which of the four components (level, trend, seasonality, noise) seem to be present in this series?


#### Question CH2-Q3 Answer:

```{r pressure, echo=FALSE}
plot(pressure)
```

## Chapter 3:

Performance Evaluation (Pages 67-68) #2 and 3

### Question 2.

Forecasting Shampoo Sales: The file ShampooSales.xls contains data on the monthly sales of a certain shampoo over a three-year period.
If the goal is forecasting sales in future months, which of the following steps should be taken?
(choose one or more below).

-   [ ] Partition the data into training and validation periods.
-   [ ] Examine time plots of the series and of model forecasts only for the training period.
-   [ ] Look at MAPE and RMSE values for the training period.
-   [ ] Look at MAPE and RMSE values for the validation period.
-   [ ] Compute naïve forecasts. Page 1 of 2


#### Question CH3-Q2 Answer:

### Question 3.

Performance on Training and Validation Data: Two different models were fit to the same time series.
The first 100 time periods were used for the training period and the last 12 periods were treated as a validation period.
Assume that both models make sense practically and fit the data reasonably well.
Below are the RMSE values for each of the models:

| Model   | Training Period | Validation Period |
|:--------|:---------------:|:-----------------:|
| Model A |       543       |        690        |
| Model B |       669       |        675        |

a)  Which model appears more useful for retrospectively describing the different components of this time series?
    Why?

b)  Which model appears to be more useful for forecasting purposes?
    Why?
    
    
#### Question CH3-Q3 Answer:
