---
title: "ADS-506 Mod 3"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
defaultW <- getOption("warn") 

options(warn = -1) 

```

## Module 3 Assignment Exercises

The assignment for this module is a mixture of programming and written
work. Complete this assignment in R Markdown. You will need to include
the question and number that you are answering within your submitted
assignment. For programming answers using R, answers should be written
in **R Markdown and 'knitted' to a Word/PDF file.**

-   Submit the written responses as a Word/PDF document. Answers must be
    clear and explicit.\
-   Submit the answer code as a PDF document. You can publish R code and
    output in R Markdown and convert it to a PDF file.

**Textbook Exercises (Pages 113-116 & 141)**\
**5.8**\
Forecasting Australian Wine Sales : Figure 5.13 shows time plots of
monthly sales of six types of Australian wines (red, rose, sweet white,
dry white, sparkling, and fortified) for 1980-1994. Data available in
AustralianWines.xls. The units are thousands of liters. You are hired to
obtain short-term forecasts (2-3 months ahead) for each of the six
series, and this task will be repeated every month

b)  Fortified wine has the largest market share of the six types of
    wine. You are asked to focus on fortified wine sales alone and
    produce as accurate a forecast as possible for the next two months

```{r Data ETL, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# # Enter R code here
library(fpp2)
library(zoo)
library(readr)

set.seed(506)

# Load Data
AustralianWines <- read_csv("Data/AustralianWines.csv",
                            col_types = cols(Month = col_date(format = "%b-%y"),
                                             Fortified = col_double(),
                                             Red = col_double(), 
                                             Rose = col_double(),
                                             sparkling = col_double(), 
                                             `Sweet white` = col_double(),
                                             `Dry white` = col_double()))


# Create Time Series Object
Fortified <- ts(AustralianWines$Fortified, start = c(1980, 1), frequency = 12)

# Training and test partitions
Fort.train <- window(Fortified, end = c(1993, 12))
Fort.test <- window(Fortified, start = c(1994, 1))


# Holt-Winter’s exponential smoothing (with multiplicative seasonality) to sales.
Fort.HWES.model <- ets(Fort.train, model = "ZZM")
Fort.HWES.pred <- forecast(Fort.HWES.model, h = 3)

accuracy(Fort.HWES.model)
accuracy(Fort.HWES.pred)


```

```{r Plot 3 Month Forecast, message=FALSE, warning=FALSE}
# Plot the model forecast.
autoplot(window(Fort.train, start = c(1993, 1)), series = 'Training') +
  autolayer(window(Fort.test, end = c(1994, 3)), series = 'Actual') +
  autolayer(Fort.HWES.pred, series = 'Prediction', alpha = .4) +
  theme_classic() 

```

c)  Create a time plot of the residuals from the Holt-Winter's
    exponential smoothing.

```{r Residual Plots}
checkresiduals(Fort.HWES.model)
```

\newpage

**5.9**\
Natural Gas Sales : Figure 5.14 shows a time plot of quarterly natural
gas sales (in billions of BTU's) of a certain company, over a period of
4 years. The company's analyst is asked to use a moving average model to
forecast sales in Winter 2005.

a)  Reproduce the time plot with the overlaying MA(4) line.

```{r}
#dataset comes from
# https://fred.stlouisfed.org/series/NATURALGASD11


# Load Data
NaturalGas <- read_csv("Data/NaturalGas.csv", 
                       col_types = cols(Quarter = col_date(format = "%m/%d/%Y"), 
                                        NaturalGas = col_double()))


# Create Time Series Object
gas <- ts(NaturalGas$NaturalGas, start = c(2001, 1), frequency = 4)


# Create rolling moving average.
gas.ma <- rollmean(gas, k = 4, align = "right")


autoplot(gas, series = "Actual", color = "blue") +
  autolayer(gas.ma , series = "MovingAvg",color = "black") +
  geom_point(color = "blue") +
  theme_classic() +
  labs(x = "Season",
     y = "Billion BTU") +
  theme(axis.title.y = element_text(face="bold"),
        axis.title.x = element_text(face="bold"),
        axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5,
                                   hjust=1),
        legend.position = "top", 
        legend.direction = "horizontal") +
  zoo::scale_x_yearqtr(format = '%Y Q%q') 
```

b)  What can we learn about the series from the MA line? 

**Answer** There is a declining trend in gas sales during this period
    
c)  Run a moving average forecaster with adequate season length. Are
    forecasts generated by this method expected to over-forecast,
    under-forecast, or accurately forecast actual sales? Why?
    
**Answer** We end up over-forecasting actual sales. This seems like a shock event happens in the winter of 2005 and early 2006 that is unforeseeable in the data. This could be considered an outlier event.
    
    
```{r}

# Test Train Partitioning
gas.train <- window(gas, end = c(2005, 3))
gas.test <- window(gas, start = c(2005, 4), end = c(2006, 2))

# Create rolling moving average.
gas.ma.model <- rollmean(gas.train, k = 4, align = "right")
gas.ma.pred <- forecast(gas.ma.model, h = 3)


# Plot the model forecast.
autoplot(gas.train, series = 'Training') +
  autolayer(gas.test, series = 'Actual') +
  autolayer(gas.ma.pred, series = 'Prediction', alpha = .4) +
  theme_classic()

```
    
    

\newpage

**6.6** Forecasting Australian Wine Sales : Figure 6.26 shows time plots
of monthly sales of six types of Australian wines (red, rose, sweet
white, dry white, sparkling, and fortified) for 1980-1994. The data is
available in AustralianWines.xls. The units are thousands of liters. You
are hired to obtain short-term forecasts (2-3 months ahead) for each of
the six series, and this task will be repeated monthly

b)  Fortified wine has the largest market share of the six types of wine
    considered. You are asked to focus on fortified wine sales alone and
    produce as accurate as possible forecasts for the next 2 months.\ **Answer** See Below

-   Start by partitioning the data using the period until December 1993
    as the training period.\ **Answer** See Below
-   Fit a regression model to sales with a linear trend and
    seasonality.\ **Answer** See Below

i.  Create the "actual vs. forecast" plot. What can you say about model
    fit?
    **Answer** The model predict pretty well looking at the "actual vs. forecast" plot we see the actual sales is within the prediction intervals of the model.
ii. Use the regression model to forecast sales in January and February
    1994
    **Answer** Forecasted Sales for January are 919. Forecasted sales for February are 1270.

```{r}

# Create Time Series Object
Fortified <- ts(AustralianWines$Fortified, start = c(1980, 1), frequency = 12)

# Training and test partitions
Fort.train <- window(Fortified, end = c(1993, 12))
Fort.test <- window(Fortified, start = c(1994, 1), end = c(1994,2))


# Fit a regression model with trend and seasonality.
fort.lm.model <- tslm(Fort.train ~ trend + season, Fort.train)
fort.lm.pred <- forecast(fort.lm.model, h = 2)

fort.lm.pred

# Plot Results
autoplot(window(Fort.train, start = c(1993, 1)), series = "Training") +
  autolayer(Fort.test, series = 'Actual') +
  autolayer(fort.lm.pred, series="Prediction", alpha = .4) +
  xlab("Date") + 
  ylab("Monthly Sales") +
  ggtitle("Forecasts for Monthly Sales") +
  theme_classic() +
  guides(colour=guide_legend(title="Forecast"))



```
